---
title: "SAT 17-18 Score Report"
author: Salman Khan
affiliation: RISE Network- East Hartford Public Schools
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    fig_caption: true
    number_sections: true
    toc: true
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(knitr)
data = read.csv("~/Data Projects/216854_SAT_20180504_100251.csv", header = TRUE, sep = ",")
```

#Executive Summary 
The following report provides a high level overview of district and school level performance on the SAT assessment. The purpose of the report is to provide stakeholders, leaders, and community members with several points of analysis regarding high school academic performance of Sample Public Schools. 

#Background
The College Board has established "college and career readiness" benchmark for the two sections of the new SAT. Meeting the benchmark means that students are on track for success beyond high school.  The target benchmark for Evidence Based Reading and Writing is 480, while the target for Math is 530. The College Board predicts that students who meet these benchmarks would earn a grade of a C or better in a first-semester college course. 

# Understanding the data
There are `r (nrow(data)-1)` students that took the SAT 11. These students come from `r n_distinct(data$AI_NAME)` unique schools.

##Demographic Data
| Gender | Number of Students |
|:--|:--:|
| Male  | `r sum(data$SEX == "M")`
| Female | `r sum(data$SEX == "F")`


| Ethnicity Categories | Number of Students |
|:--|:--:|
| Cuban  | `r sum(data$RACE_ETH_CUBAN == "Y")`
| Mexican | `r sum(data$RACE_ETH_MEXICAN == "Y")`
| Puerto Rican | `r sum(data$RACE_ETH_PUERTORICAN == "Y")`
| Hispanic/Latino | `r sum(data$RACE_ETH_HISP_LAT == "Y")`
| Non-Hispanic | `r sum(data$RACE_ETH_PUERTORICAN == "Y")`

| Race Categories | Number of Students |
|:--|:--:|
| American Indian/Alaska Native | `r sum(data$RACE_ETH_INDIAN_ALASKAN == "Y")`
| Asian | `r sum(data$RACE_ETH_ASIAN == "Y")`
| African American | `r sum(data$RACE_ETH_AFRICANAMERICAN == "Y")`
| White | `r sum(data$RACE_ETH_WHITE == "Y")`
| Native Hawaiian/Pacific Islander | `r sum(data$RACE_ETH_HAWAIIAN_PI == "Y")`
| Other | `r sum(data$RACE_ETH_OTHER == "Y")`


#Summary Statistics 
The lowest score was a **`r min(data$LATEST_SAT_TOTAL)`** and the highest score was a **`r max(data$LATEST_SAT_TOTAL)`**. 

##Students Meeting College-Readiness Benchmarks
| Section | Number of Students | Percent of Total Students |
|:--|:--:|:--:|
| EBRW | `r EBRW <- nrow(data[data$LATEST_SAT_EBRW >= 480, ])` `r EBRW` | `r round((EBRW/ (nrow(data)-1))*100, 2)`%
| Math | `r MATH_SECTION <- nrow(data[data$LATEST_SAT_MATH_SECTION >= 530, ])` `r MATH_SECTION` | `r round((MATH_SECTION/ (nrow(data)-1))*100, 2)`%
| Both | `r Both_CR <- nrow(data[data$LATEST_SAT_MATH_SECTION >= 530 & data$LATEST_SAT_EBRW >= 480, ])` `r Both_CR` | `r round((Both_CR/ (nrow(data)-1))*100, 2)`%

##Average Total Scores

| Total | EBRW | Math |
|:--:|:--:|:--:|
| `r round(mean(data$LATEST_SAT_TOTAL), 2)` | `r round(mean(data$LATEST_SAT_EBRW), 2)` | `r round(mean(data$LATEST_SAT_MATH_SECTION), 2)`

###Average Total Scores Over Time
|  | Total | EBRW | Math |
|:--:|:--:|:--:|:--:|
| SAT 11 | `r round(mean(data$LATEST_SAT_TOTAL), 2)` | `r round(mean(data$LATEST_SAT_EBRW), 2)` | `r round(mean(data$LATEST_SAT_MATH_SECTION), 2)`


##Average Sub-Section Scores

| Reading | Writing & Language | Math Section |
|:--:|:--:|:--:|
| `r round(mean(data$LATEST_SAT_READING), 2)` | `r round(mean(data$LATEST_SAT_WRIT_LANG), 2)` | `r round(mean(data$LATEST_SAT_MATH_TEST), 2)`

##Average Cross-Section Scores

| Science Cross | History Cross | 
|:--:|:--:|
| `r round(mean(data$LATEST_SAT_SCI_CROSS), 2)` | `r round(mean(data$LATEST_SAT_HIST_SOCST_CROSS), 2)` |

##Average EBRW Question-Type Scores

| Words in Context | Command of Evidence | Expression of Ideas | Standard English Conventions |
|:------:|:------:|:------:|:-----------:|
| `r round(mean(data$LATEST_SAT_WORDS_CONTEXT), 2)` | `r round(mean(data$LATEST_SAT_COMM_EVIDENCE), 2)` | `r round(mean(data$LATEST_SAT_EXPR_IDEAS), 2)` | `r round(mean(data$LATEST_SAT_ENG_CONVENT), 2)` 

##Average Math Question-Type Scores

| Heart of Algebra | Advanced Mathematics | Problem Solving & Data Analysis |
|:--:|:--:|:--:|
| `r round(mean(data$LATEST_SAT_HEART_ALGEBRA), 2)` | `r round(mean(data$LATEST_SAT_ADV_MATH), 2)` | `r round(mean(data$LATEST_SAT_PROBSLV_DATA), 2)`


#Potential Graphs

##Plot of EBRW by Math:

```{r, warning = FALSE, message = FALSE}
  ggplot(data, aes(x = data$LATEST_SAT_EBRW, y = data$LATEST_SAT_MATH_SECTION)) +
  geom_point() +
  geom_smooth() +
  labs(x = "EBRW", y = "Math", title = "Relationship between EBRW and Math Sections")
```

##Histogram: 

```{r}
ggplot(data, aes(x = data$LATEST_SAT_TOTAL)) + 
  geom_histogram(binwidth = 200, color = "mediumblue", fill = "steelblue2") +
  labs(x = "Total SAT Scores", y="Frequency", title="Distribution of Total SAT Scores") +
  geom_vline(aes(xintercept=mean(data$LATEST_SAT_TOTAL)),
            color="goldenrod2", linetype="dashed", size=1) + 
  theme_minimal()

```


